Prognostic and biological relevance of collagen-related genes in prostate cancer

### 2024-01-31 report/markdown/manuscript_supplement.Rmd
Confusion matrix heat maps, PIRADS3

Transcriptome cohorts for analysis in the current report were selected from studies deposited at cBioportal and Gene Expression Omnibus (GEO) with the following criteria: availability of information on Gleason scoring, biochemical relapse and biochemical relapse-free survival, and expression data for `r nrow(globals$genes_interest)` collagen-related genes investigated also in the proteomic part of the project (__Supplementary Table S\@ref(tab:tab-cohorts)__ and __S\@ref(tab:tab-genes)__).

@@

Gene expression levels were transformed with the $log_2(expression)$ and $log_2(count + 1)$ function for the microarray and RNA sequencing data sets prior to further analyses. 
The _MCP Counter_ and _xCell_ estimates of non-malignant cell content in cancer samples were computed with the `deconvolute()` function from the _immunedeconv_ package. 
Gene signatures of the [Reactome pathways](https://reactome.org/) were obtained from the [MSig database](https://www.gsea-msigdb.org/gsea/msigdb/index.jsp), version 7.5.1. 
ssGSEA scores were computed with the `calculate()` function from the [_gseaTools_](https://github.com/PiotrTymoszuk/gseaTools) package.

@@

flextable::flextable(suppl_paper_tbl$clinic) %>% 
  set_widths(c(2.8, 3.5, 4, 4, 2.8, 2.8)) %>% 
  merge_v(1) %>% 
  footnote(1, 2, 
           value = as_paragraph('PSA: prostate-specific antigen.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 5:6, 
           value = as_paragraph('Qualitative variables: \u03C7\u00B2 test with Cramer V effect size statistic. Numeric variables: Mann-Whitney test with r effect size statistic. P values corrected for multiple testing with the false discovery rate.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

@@

```{r fig-clust-tumor, fig.width = figur::convert(suppl_paper_fig$clust_tumor, to = 'in')$w, fig.height = figur::convert(suppl_paper_fig$clust_tumor, to = 'in')$h, fig.cap = 'Gleason score and pathological tumor stage in the collagen clusters.'}

@@

__Supplementary Figure S\@ref(fig:fig-clust-tumor). Gleason score and pathological tumor stage in the collagen clusters.__

_Differences in distribution of Gleason scores (A, 5 - 6, 7 and $\geq$ 8) and pathological tumor stages (B) between the collagen clusters were investigated by $\chi^2$ test with Cramer V effect size statistic. P values were corrected for multiple testing with the false discovery rate method. Percentages of samples assigned to the Gleason score strata or tumor stages within each cluster are shown in stack plots. Effect sizes and p values are displayed in the plot facets. Numbers of complete observations in the clusters are indicated in the Y axes. Full clinical and pathological characteristic of the collagen clusters is shown in Supplementary Table S\@ref(tab:tab-clinic)._

>>>>

Transcriptome cohorts for analysis in the current report were selected from studies deposited at cBioportal and Gene Expression Omnibus (GEO) with the following criteria: availability of information on Gleason scoring or ISUP risk strata (International Society of Urological Pathology), biochemical relapse and biochemical relapse-free survival, and expression data for `r nrow(globals$genes_interest)` collagen-related genes investigated also in the proteomic part of the project (__Supplementary Table S\@ref(tab:tab-cohorts)__ and __S\@ref(tab:tab-genes)__).

@@

Gene expression levels were transformed with the $log_2(expression)$ and $log_2(count + 1)$ function for the microarray and RNA sequencing data sets prior to further analyses. 
The _MCP Counter_ and _xCell_ estimates of non-malignant cell content in cancer samples were computed with the `deconvolute()` function from the _immunedeconv_ package. 
Gene signatures of the [Reactome pathways](https://reactome.org/) were obtained from the [MSig database](https://www.gsea-msigdb.org/gsea/msigdb/index.jsp), version 7.5.1. 
ssGSEA scores were computed with the `calculate()` function from the [_gseaTools_](https://github.com/PiotrTymoszuk/gseaTools) package. 
ISUP strata were defined based on Gleason Score information as follows: ISUP1: Gleason score 5 - 6, ISUP2: Gleason score 7, ISUP3+: Gleason score $\geq$ 8.

@@

flextable::flextable(suppl_paper_tbl$clinic) %>% 
  set_widths(c(2.8, 3.5, 4, 4, 2.8, 2.8)) %>% 
  merge_v(1) %>% 
  footnote(1, 2, 
           value = as_paragraph('PSA: prostate-specific antigen; ISUP: International Society of Urological Pathology risk stratification system.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 5:6, 
           value = as_paragraph('Qualitative variables: \u03C7\u00B2 test with Cramer V effect size statistic. Numeric variables: Mann-Whitney test with r effect size statistic. P values corrected for multiple testing with the false discovery rate.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

@@

```{r fig-clust-tumor, fig.width = figur::convert(suppl_paper_fig$clust_tumor, to = 'in')$w, fig.height = figur::convert(suppl_paper_fig$clust_tumor, to = 'in')$h, fig.cap = 'ISUP and pathological tumor stage in the collagen clusters.'}

@@

__Supplementary Figure S\@ref(fig:fig-clust-tumor). ISUP and pathological tumor stage in the collagen clusters.__

_Differences in distribution of ISUP risk strata (International Society of Urological Pathology: ISUP1, ISUP2 and ISUP3+) and pathological tumor stages (B) between the collagen clusters were investigated by $\chi^2$ test with Cramer V effect size statistic. P values were corrected for multiple testing with the false discovery rate method. Percentages of samples assigned to the ISUP strata or tumor stages within each cluster are shown in stack plots. Effect sizes and p values are displayed in the plot facets. Numbers of complete observations in the clusters are indicated in the Y axes. Full clinical and pathological characteristic of the collagen clusters is shown in Supplementary Table S\@ref(tab:tab-clinic)._
